---
title: "ma615 1004"
author: "Luyin Fu"
date: "10/3/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(magrittr)
library(kableExtra)
library(ggplot2)              
library(esquisse)             
library(kableExtra)
library(tidyr)
library(stringr)
library(dplyr)
```

## question1

### How many continents are included in the data set?  
``` {r}
library(gapminder)
data("gapminder")

a1 <- length(unique(gapminder$continent))
a2 <- length(unique(gapminder$country))
```
\qquad the number of continents in the data is `r a1`  

### How many countrys are included? How many countries per continent?

\qquad `r a2` countries are involved in the gapminder dataset  
```{r}
A <- gapminder %>% group_by(continent) %>%  
  summarise(countries =length(unique(country)))

kable(A, format = "latex", booktabs=TRUE,
      col.names = c("continent", "number"),
      caption = "the number of countries involved") %>% kable_styling(latex_options = "HOLD_position")
```
### Using the gapminder data, produce a report showing the continents in the dataset, total population per continent, and GDP per capita.

```{r}
G1 = gapminder %>% group_by(year,continent) %>% 
  summarise(tol_POP = sum(as.numeric(pop)), GDP_perCap = sum(pop*gdpPercap)/tol_POP) %>% 
  arrange(continent)
colnames(G1)=c("Year","Continent","Total Population","GDP Per Capita")
g1 = cbind(G1[1:12,],G1[13:24,],G1[25:36,])
g2 = cbind(G1[37:48,],G1[49:60,])

kable(g1[,c(1,3,4,7,8,11,12)], caption = "Total Population and GDP per Capita", align = "c", booktab =T,format = "latex") %>% kable_styling(latex_options = c("HOLD_position","scale_down")) %>% add_header_above(c("Year"=1,"Africa"=2,"American"=2,"Asia"=2))

kable(g2[,c(1,3,4,7,8)], caption = "Total Population and GDP per Capita", align = "c", booktab =T,format = "latex") %>% kable_styling(latex_options = c("HOLD_position")) %>% add_header_above(c("Year"=1,"Europe"=2,"Oceania"=2))

```

### Produce a well-labeled table that summarizes GDP per capita for the countries in each continent, contrasting the years 1952 and 2007.
```{r}
GDP_1952 = gapminder %>% filter(year == "1952") %>% 
  group_by(continent) %>% 
  summarise(mean = mean(gdpPercap), min=min(gdpPercap), max=max(gdpPercap), sd=sd(gdpPercap)) %>% 
  arrange(mean)
kable(GDP_1952, caption = " GDP per Capita Summary 1952", align = "c", booktab =TRUE, format = "latex") %>% 
  kable_styling(latex_options = c("HOLD_position"))

GDP_2007 = gapminder %>% filter(year == "2007") %>% 
  group_by(continent) %>% 
  summarise(mean = mean(gdpPercap), min=min(gdpPercap), max=max(gdpPercap), sd=sd(gdpPercap)) %>% 
  arrange(mean)
kable(GDP_2007, caption = " GDP per Capita Summary 2007", align = "c", booktab =TRUE, format = "latex") %>% 
  kable_styling(latex_options = c("HOLD_position","scale_down"))
```

### Product a plot that summarizes the same data as the table. There should be two plots per continent.
```{r}
GDP_both <- gapminder %>% filter(year == "1952" |year == "2007") %>% 
  group_by(year, continent) %>% 
  mutate(mean = mean(gdpPercap), min=min(gdpPercap), max=max(gdpPercap), sd=sd(gdpPercap))  %>%
  select(country, continent, year, gdpPercap, mean, sd) 

ggplot(GDP_both, aes(x = country, y = gdpPercap))+
  geom_jitter()+
  facet_grid(continent ~ year)
```

### Which countries in the dataset have had periods of negative population growth? Illustrate your answer with a table or plot.

```{r}


```

### Which countries in the dataset have had the highest rate of growth in per capita GDP? Illustrate your answer with a table or plot.
```{r}
gapminder %<>% mutate(log_growth=log(gdpPercap))
G3 = gapminder %>% group_by(country) %>% 
  summarise(highest_growth_rate = max(diff(log_growth))) %>% 
  arrange(desc(highest_growth_rate)) %>%
  slice(1:10)
ggplot(G3, aes(x=country, y = highest_growth_rate, fill=country))+
  geom_bar(stat="identity")+
  theme_bw()+
  ggtitle("top ten yearly growth rate countries ever had")
```

## question2

### There are four possible gender combinations for the first two Children. Product a plot the contracts the frequency of these four combinations.
```{r}
data(Fertility, package="AER")

l <- length(Fertility$gender1)
F1 <- Fertility %>% group_by(gender1,gender2) %>% 
  summarise(frequency=n()/l)
kable(F1, format = "latex", booktabs=TRUE, digits = 3,
      col.names = c("gender1", "gender2", "frequency"),
      caption = "frequency of first two baby gender") %>% kable_styling(latex_options = "HOLD_position")

```



### Are the frequencies different for women in their 20s and wemen who are older than 29?
```{r}
l1 <- sum(Fertility$age >29)
F2 <- Fertility %>% group_by(gender1,gender2) %>% 
  filter(age >29) %>%
  summarise(frequency=n()/l1)

l2 <- l-l1
F3 <- Fertility %>% group_by(gender1,gender2) %>% 
  filter(age <= 29) %>%
  summarise(frequency=n()/l2)
F4 <- cbind(F2,F3)
colnames(F4) <- rep(c('baby1 gender', 'baby2 gender', 'frequency'),2)
kable(F4, format = "latex", booktabs=TRUE, 
      caption = "frequency comparison") %>%
  kable_styling(latex_options = "HOLD_position") %>%
  add_header_above(c("age before 30"=3, "age above 30"=3))
```
the frequencies are pretty similar

### Produce a plot that contrasts the frequency of having more than two children by race and ethnicity.
```{r}
Fertility <- mutate(Fertility, race=ifelse(afam=="yes","afam",
                              ifelse(hispanic=="yes", "hispanic",
                               ifelse(other=="yes", "other", NA))))
F5 <- drop_na(Fertility)
#l3 <- length(F5$age)
Freq_2more <- F5 %>% group_by(race) %>% 
  summarise(frequency = sum(morekids=="yes")/n())
ggplot(Freq_2more, aes(x=race, y=frequency))+
  geom_bar(stat="identity",fill="steelblue")+
  theme_minimal()+
  ggtitle("frequency of having more than two children")
```


## question 3

### How many times does the letter “e” occur in mtcars rownames?  
### How many cars in mtcars have the brand Merc?
### How many cars in mpg have the brand(“manufacturer” in mpg) Merc?
``` {r}
data("mtcars")
data("mpg")

mtcars1 <- mutate(mtcars,carnames=rownames(mtcars))
m1 <- sum(str_detect(mtcars1$carnames,"e"))

m2 <- sum(str_detect(mtcars1$carnames,"Merc"))

mpg_brand <- data.frame(brand=unique(mpg$manufacturer))
m3 <- sum(str_detect(mpg_brand$brand,"merc"))
```
there are `r m1` cars in mtcars with a 'e' in their names  
there are `r m2` cars in mtcars that have the brand Merc  
there are `r m3` cars in mpg that have the brand Merc

### Contrast the mileage data for Merc cars as reported in mtcars and mpg.
```{r}
M1 <- mtcars1 %>% select(mpg, carnames) %>% 
  pivot_wider(names_from = carnames, values_from = mpg) %>%
  select(contains("Merc")) 
M2 <- mpg %>% filter(manufacturer=="mercury") %>% 
  select(manufacturer, hwy) 
kable(M1, format = "latex", booktabs=TRUE, 
      caption = "Merc cars mileage data (miles per gallon)") %>%
  kable_styling(latex_options = "HOLD_position") 

```

## question 4
```{r}
library(babynames)
data("babynames")
```
### Draw a sample of 500,000 rows from the babynames data
This is a glimpse of a sample of the babyname data
``` {r}
bbname_sample= sample_n(babynames, 500000, replace = FALSE)
glimpse(bbname_sample)
```

### Produce a tabble that displays the five most popular boy names and girl names in the years 1880,1920, 1960, 2000.
The 5 most popular baby names for both gender is presented in the table.
```{r}
babynames1880 <- filter(babynames, year=="1880")
babynames1920 <- filter(babynames, year=="1920")
babynames2000 <- filter(babynames, year=="2000")

B1 <- babynames1880 %>% group_by(sex) %>%
  arrange(desc(n), .by_group= TRUE) %>%
  slice(1:5) %>% select(sex, name)
B2<- babynames1920 %>% group_by(sex) %>%
  arrange(desc(n), .by_group= TRUE) %>%
  slice(1:5) %>% select(sex, name)
B3<- babynames2000 %>% group_by(sex) %>%
  arrange(desc(n), .by_group= TRUE) %>%
  slice(1:5) %>% select(sex, name)
B <- cbind(B1, B2, B3)

colnames(B) <- rep(c('sex', 'name'),3)
kable(B, format = "latex", booktabs=TRUE,
      caption = "top 5 baby names")%>% kable_styling(latex_options = "HOLD_position") %>% add_header_above(c("1880"=2, "1920"=2,"2000"=2))
```

### What names overlap boys and girls?
```{r}
B4 <- babynames %>% select(name,sex) %>%
  group_by(name) %>% 
  summarise(gender_neutral=n_distinct(sex)) %>%
  filter(gender_neutral==2)
BB <- sample_n(B4, size = 10, replace = FALSE)
```
there are `r length(B4$name)` names used both for boys and girls, here are some of the names: `r BB$name`

### What names were used in the 19th century but have not been used in the 21sth century?

```{r}
b1 <-  unique(filter(babynames, year<1900)$name)
b2 <-  unique(filter(babynames, year>2000)$name)
B5 <- setdiff(b1,b2)
```
There are `r length(B5)` names used in 19th century but have not been used in the 21sth century, such as `r B5[1:5]`  


### Produce a chart that shows the relative frequency of the names “Donald”, “Hilary”, “Hillary”, “Joe”, “Barrack”, over the years 1880 through 2017.

```{r}
B6 <- babynames %>% 
  filter( year>=1880 & year<= 2017 & name==c("Donald", "Hilary", "Hillary", "Joe", "Barack")) %>%
  group_by(name) %>% summarise(total_count=sum(n))
B6 %<>% mutate(frequency=round(total_count/sum(total_count),6)) %>%
  select(-total_count)

ggplot(B6, aes(x=name, y=frequency, fill=name))+
  geom_bar(stat="identity")+
  theme_minimal()+
  ggtitle("relative frequency of selected names 1880-2017")
```
